### 2012-01-11 article.tex
More CI

\documentclass[a4paper,11pt]{article}
\usepackage[utf8x]{inputenc}

%opening
\title{SCM plan for a troubling company}
\author{Jonas Nolte, dt07jn1\\
David Everlöf, dt08de2\\
Tobias Petersen, et07tp3}

\begin{document}

\maketitle

\begin{abstract}
Many companies, especially the small ones, don't see the benefits of SCM in their company and decides to not use it.
This company however have now seen the light and need help to implement it.
This report goes through some aspects in this implementation and focus on a relatively small software company of 40-50 employees.
We analyze some of their problems and come up with solutions in the form of a configuration plan.
\end{abstract}
\newpage

\tableofcontents
\newpage

\section{Problem}

\subsection{The Company}

The company that should have SCM implemented will be referred to as ''The Problem Company'' throughout the report.
It is a small company, 12 employees which have grown in the past few years and intends to keep growing in the future.
The current size is about 40-50 employees and develops software to a local cellphone company.

The company makes a new release to every new phone the local company produces which means they have a new release every 6 months.
They also need to maintain old cellphones since the local company updates these and The Problem Company needs to come up with new software then.

The company are programming Java and are using Eclipse as their tool.

\subsection{The Software}

The software is the messaging portion of the phone related to recieving and sending textmessages in various forms over \emph{IM-networks} and  \emph{sms}. There are plans to integrate facebook as well.
The rapidly changing standards and lack of documentation from the IM-vendors make this a time consuming job and the software needs to be updated on a regular basis to support the new protocols.
In the beginning there was only one version of the application that was supported but now they have implemented some new features that old phones can't handle but these versions still needs updates to keep supporting the old features.
The software is developed in modules with small groups working on separate modules such as the UI and the different backends for various protocols.
The separation is ok the modules can be developed separately most of the time but occasionally synchronisation is needed between releases for API changes or new features affecting the backends.
If synchronisation is needed for API changes and the like it's solved in an ad hoc fashion by copying files on thumbdrives or internal emails.
The documentation is modified by whoever changes the API as microsoft doc files on a central fileserver.
When a release is needed the different modules are copied to a central computer and some persons from each team try to merge together these files by hand until they can produce a build that works.

\subsection{Documentation}
The documentation is modified by whoever changes the API as microsoft doc files on a central fileserver.
Conflicts are avoided by keeping those files open if you're trying to change them.
This has worked quite well but the recent size increase makes problems with people waiting on these files and forgetting to close them a problem.

\subsection{The build}

\subsection{Perceived problems}

\begin{itemize}
\item They spend more and more time on resolving merge conflicts.
\item Previously solved bugs finds it’s way back into the code again.
\item Integrating parts of the system takes longer and longer time.
\item Problem to deliver the product to the customer within deadlines.
\item When customers have problem with previous releases the company can't fix this easily.
\item If the company keeps growing, the problems will grow as well and at some point become impossible to handle.
\item Files and work are getting lost as there is no central repository of them with backups.
\end{itemize}

\section{Configuration Plan}

The term software configuration management is for many companies unknown and they have no clue on how to describe it.
But in fact they might use some of the concepts without even knowing it.
Companies can come up with their own solutions and integrate them into the system and can be either good or bad.
When we are about to write the configuration plan we will have to tailor SCM to this company and have in mind things like, which programming language the company is using and which tools are used.

The configuration plan can be written in two different approaches. \cite{SCMPLAN} It can either be done before the implementation of SCM in a big bang approach or we can use a more iterative method and complete the plan meanwhile the project goes on.
Since we are new to the role as a configuration manager and don’t have a precise definition on the company and its details it might be too hard to go with the iterative approach.
This way we can start with the most important parts of the SCM plan in the beginning

To write a good configuration plan can be hard, it is different for every company and will determinate the future of SCM in the future.
A good SCM plan can rise a sinking company as well sink a well working.
In an article written by Bounds and Dart they describe the SCM plan as one of the three most important keys to successfully attain a SCM solution \cite{BoundsDart}.
It is possible to write one all by yourself as a configuration manager but with our small experience in the area we will take help from someone with more skills.
We will also have good contact with the product manager to make this plan as good as possible.

A SCM plan is a good checklist for the configuration manager to check whether all is kept in mind for the SCM implementation or something is missed. When all the checkpoints are completed we have a completed SCM plan to work upon and we know we have thought of all the details. Once the configuration plan is completed it’s not finished. Changes are often made to the project which will have impact on the plan, we therefore need to change the configuration plan as well so it is always updated. This is one of the reasons why we need to have the configuration plan as a configuration item (CI) which we will describe later on in the paper.

A complete configuration plan will cover everything to implement SCM to a specific company and is a very large document. Here in this report we will therefore only discuss a few aspects of it which we decide are the most important and interesting. The SCM plan will discuss how to solve The Problem Companiy's problems.

\section{Configuration Identification}

\subsection{Background}
The identification part of the plan contains useful information about how objects should be treated. One must decide whether or not the object should be under version control. When this should be decided one should ask the question ``Is it necessary to have this object under version control in order to be able to build the complete, working release?''. If the answer to that question is yes, then the object should be configuration item (an item under version control).
Usually, the major part of the object should be under version control. What usually shouldn't be under version control is items such as PM, mails, announcements, memos.

Also labeling and numbering scheme must be decided. The numbering scheme for forms must also be specified, such as for CCB documents etc.

In the identification process the different baselines must be identified. And more specifically who creates them, authorizes and verifies them, as well as the contents of them.\cite{cagan}

\subsection{This software}

The configuration identification for the software part is quite straightforward where every module is it's own configuration item and can be subdivided in the files as the next step of the hierarchy.
For the files the filenames and their directories provides a natural naming and an SCM-tool makes this easy to keep track of. The various modules are then combined and this can be kept in the tool as well or be managed manually.
As the documentation actually is tied to the code and it's versions it should, if possible be controlled in the same system to keep it in sync with the code.
The next higher level of configutration control needed is the release level where the build environment and the actual produced binaries together with the released documentation.
This also needs to be documented and controlled in some way as the released models are supported for some years at least.

\subsection{Other items}

As the delivered product is in form of binaries and documentation for inclusion in a larger product there is not much other releasables to control.
However to structure the change process and design of new features design and specification documents are created for communication between customer and the company as well as between internal designers and the coders. These needs to be versioned and saved as well to be able to verify that the product performs according to the specifications.

\section{Configuration Control}

The main idea behind configuration control is to achieve control of changes, what impact they'll have and their costs.

Since this company is relatively small, the extent of CCB and it's usage will be a somewhat limited in contrary to the general idea of a CCB.
The CCB consist of the normal lineup with: chairperson, secretary, members and specialists \cite{daniels}. We will select the project manager
to be the chairman and the secretary will be the configuration manager in the company. The project manager will appoint appropriate people to be
members of the CCB and specilists will be contacted on a ``when needed''-basis.

Not every change will be handled by the CCB in our case

Hur stor ccb?
Hur rigorös CR kontroll?

\subsection{Branching Pattern}

One of The Problem Company's problems is that they spend more and more time on resolving merge conflicts.
This could be the problem caused by a wrong branch pattern.

\subsection{Bill of Material}

The company releases a new product every 6 months which results we have several different products to maintain. A problem is that when a customer finds a bug or wants a new feature in a previous release The Problem Company can't reproduce this system or it's very hard. We will need something so we both know exactly what system the customer is using and so the programmer who will fix this problem easy can reconstruct the system to work on. A way to do this is to use a Bill of Material (BoM). When we release a product we then save information about the release such as which version every component is using, what hardware are we using, which features are implemented, test information etc.
When a customer has a problem with a certain release we can look at the BoM and know exactly what system the customer is using and we can start working on a fix right away.
We will also need version control of the product to get the correct version of each component in the program.
Everyone in the project must be able to find this BoM without too much effort and we need a system which makes it easy to setup a certain BoM on the computer which the programmer is using. Otherwise we loose the benefits with the BoM. The software which the company is producing are used on specific cellphones so we need to be able to reproduce the hardware somehow. This can be done by storing some cellphones but very often will the software not interact with the actual hardware and we can then focus on just software information. In the BoM we store both software information about our product but also the software the cellphone is using.

\section{Tools}
With eclipse they have several different add-ons which makes it hard for them to use another tool. An approach building upon this platform, in the from of another plug-in would therefore probably be the best solution.

\end{document}