### 2012-01-17 article.tex
Lots of new content

\documentclass[a4paper,11pt]{article}
\usepackage[utf8x]{inputenc}
\usepackage{cite}

%opening
\title{SCM plan for a troubling company}
\author{Jonas Nolte, dt07jn1\\
David Everlöf, dt08de2\\
Tobias Petersen, et07tp3}

\begin{document}

\maketitle

\begin{abstract}
Many companies, especially the small ones, don't see the benefits of SCM in their company and decides to not use it.
This company however have now seen the light and need help to implement it.
This report goes through some aspects in this implementation and focus on a relatively small software company of 40-50 employees.
We analyze some of their problems and come up with solutions in the form of a configuration plan.
\end{abstract}
\newpage

\tableofcontents
\newpage

\section{Problem}

\subsection{The Company}

The company that should have SCM implemented will be referred to as ''The
Problem Company'' throughout the report. It is a small company, 12 employees
which have grown in the past few years and intends to keep growing in the
future. The current size is about 40-50 employees and develops software to a
local cellphone company. The Problem Company is located in Lund and has no plans
on moving or expand to other cities. Therefore all who works in the project are
located in the same building, however the customer is located in a neighboring
building and needs to have good contact with them.

The company makes a new release to every new phone the local company produces
which means they have a new release every 6 months. They also need to maintain
old cellphones since the local company updates these and The Problem Company
needs to come up with new software then.

The company are programming Java and are using Eclipse as their tool. Since the
beginning they have used traditional programming instead of agile and are well
familiar and comfortable with it.

\subsection{The Software}

The software is the messaging portion of the phone related to recieving and
sending textmessages in various forms over \emph{IM-networks} and  \emph{sms}.
There are plans to integrate facebook as well. The rapidly changing standards
and lack of documentation from the IM-vendors make this a time consuming job and
the software needs to be updated on a regular basis to support the new
protocols. In the beginning there was only one version of the application that
was supported but now they have implemented some new features that old phones
can't handle but these versions still needs updates to keep supporting the old
features. The software is developed in modules with small groups working on
separate modules such as the UI and the different backends for various
protocols. The separation is ok the modules can be developed separately most of
the time but occasionally synchronisation is needed between releases for API
changes or new features affecting the backends. If synchronisation is needed for
API changes and the like it's solved in an ad hoc fashion by copying files on
thumbdrives or internal emails. The documentation is modified by whoever changes
the API as microsoft doc files on a central fileserver. When a release is needed
the different modules are copied to a central computer and some persons from
each team try to merge together these files by hand until they can produce a
build that works.

\subsection{Documentation}
The documentation is modified by whoever changes the API as microsoft doc files
on a central fileserver. Conflicts are avoided by keeping those files open if
you're trying to change them. This has worked quite well but the recent size
increase makes problems with people waiting on these files and forgetting to
close them a problem.

\subsection{The build}

\subsection{Perceived problems}

\begin{itemize}
\item They spend more and more time on resolving merge conflicts.
\item Previously solved bugs finds it’s way back into the code again.
\item Integrating parts of the system takes longer and longer time.
\item Problem to deliver the product to the customer within deadlines.
\item When customers have problem with previous releases the company can't fix this easily.
\item If the company keeps growing, the problems will grow as well and at some point become impossible to handle.
\item Files and work are getting lost as there is no central repository of them with backups.
\end{itemize}

\section{Configuration Plan}

The term software configuration management is for many companies unknown and they have no clue on how to describe it.
But in fact they might use some of the concepts without even knowing
it. Companies can come up with their own solutions and integrate them into the
system and can be either good or bad. When we are about to write the
configuration plan we will have to tailor SCM to this company and have in mind
things like, which programming language the company is using and which tools are
used.

The configuration plan can be written in two different approaches.
\cite{SCMPLAN} It can either be done before the implementation of SCM in a big
bang approach or we can use a more iterative method and complete the plan
meanwhile the project goes on. Since we are new to the role as a configuration
manager and don’t have a precise definition on the company and its details it
might be too hard to go with the iterative approach. This way we can start with
the most important parts of the SCM plan in the beginning

To write a good configuration plan can be hard, it is different for every
company and will determinate the future of SCM in the future. A good SCM plan
can rise a sinking company as well sink a well working. In an article written by
Bounds and Dart they describe the SCM plan as one of the three most important
keys to successfully attain a SCM solution \cite{BoundsDart}. It is possible to
write one all by yourself as a configuration manager but with our small
experience in the area we will take help from someone with more skills. We will
also have good contact with the product manager to make this plan as good as
possible.

A SCM plan is a good checklist for the configuration manager to check whether 
all is kept in mind for the SCM implementation or something is missed. When all
the checkpoints are completed we have a completed SCM plan to work upon and we
know we have thought of all the details. Once the configuration plan is
completed it’s not finished. Changes are often made to the project which will
have impact on the plan, we therefore need to change the configuration plan as
well so it is always updated. This is one of the reasons why we need to have the
configuration plan as a configuration item (CI) which we will describe later on
in the paper.

A complete configuration plan will cover everything to implement SCM to a 
specific company and is a very large document. Here in this report we will
therefore only discuss a few aspects of it which we decide are the most
important and interesting. The SCM plan will discuss how to solve The Problem
Companiy's problems.

\section{Configuration Identification}

\subsection{Background}
The identification part of the plan contains useful information about how
objects should be treated. One must decide whether or not the object should be
under version control. When this should be decided one should ask the question
``Is it necessary to have this object under version control in order to be able
to build the complete, working release?''. If the answer to that question is
yes, then the object should be configuration item (an item under version
control). Usually, the major part of the object should be under version control.
What usually shouldn't be under version control is items such as PM, mails,
announcements, memos.

Also labeling and numbering scheme must be decided.
The numbering scheme for forms must also be specified, such as for CCB
documents etc.

In the identification process the different baselines must be identified. And
more specifically who creates them, authorizes and verifies them, as well as the
contents of them \cite{cagan}.

\subsection{This software}

The configuration identification for the software part is quite straightforward
 where every module is it's own configuration item and can be subdivided in the
files as the next step of the hierarchy. For the files the filenames and their
directories provides a natural naming and an SCM-tool makes this easy to keep
track of. The various modules are then combined and this can be kept in the tool
as well or be managed manually. As the documentation actually is tied to the
code and it's versions it should, if possible be controlled in the same system
to keep it in sync with the code. The next higher level of configutration
control needed is the release level where the build environment and the actual
produced binaries together with the released documentation. This also needs to
be documented and controlled in some way as the released models are supported
for some years at least.

\subsection{Other items}

As the delivered product is in form of binaries and documentation for inclusion
in a larger product there is not much other releasables to control. However to
structure the change process and design of new features design and specification
documents are created for communication between customer and the company as well
as between internal designers and the coders. These needs to be versioned and
saved as well to be able to verify that the product performs according to the
specifications.

\section{Configuration Control}

\subsection{Branching Pattern}

As the company hasn't used any form of SCM-tool previously explicit branching didn't exist.
The copying of code between subgroups working on different modules however
created implicit branches whenever two version weren't in sync. This creates an
implicit feature branch model with merges at each release and sometimes in
between but not very often as those merges are hard to do and requires that all
of the changes made by the other team be merged together with the needed
features/changes. This makes the merges very hard to do when it's finally time
to do so and is the cause of the long merging periods/missed deadlines.

The sporadic merges also gives a composition model but where the choices to make
a composition of are very limited (the one we have since release or a dev. one)
leading to collaboration problems. The teams mostly needs to work against the
modules from the previous release but occasionally the need to test or work
against an intermediate version from another group exists. This is frequent
enough that the possiblility to do this is a requirement from the system.

While being able to use the developing version from another group is important
so is the need for the other group to be able to continue their work without
risk of breaking the build for the first. This can be a serious problem as even
if the group is extremely thorough with testing their changes before commit an
API change must now be synchronized with the other group wich slows development
down a lot.

The branching pattern we propose on the module level are to have each group to 
develop their module in  a developement branch and when it's time for a release
or synchronization with another team to create a branch for that. This is
described in \cite{StreamedLines} as \emph{deffered branching} pattern since
the development happens in mainline and is branched when needed for
incompatible changes. The releases can then be composed from those branches.
This pattern also avoids a too wide branching structure as the releases don't
have branches of their own and thus will minimize integration work needed when
mergin branches as the need for merges should be limited to backported
bugfixes from the mainline to the release branches.

The release branches are the groups releases versus the rest of the company and
used by the higher level of the CI hierarchy as items to compose releases of
the product from. As such they have a high need for stability and there would
need to be either a freeze of the development line prior to branching of the
release or the release branch have to be done quite some time before the
release to ensure the build is as stable as needed. The freeze necessarily
stops development of new features until the branch is made while the second
approach causes more integration between the branches to be made (as any bugs
found will need to be fixed on both branches). This decision can be made on a
release by release base depending on the current needs.

\subsubsection{Team variations}

The above describes the strategy main branches and patterns all groups should
use to facilitate working releases and collaboration between the groups. In
maintaining the development branch the groups are free to any strategy they
like inform of individual branches for each developer or feature that are then
merged with the main branch or just doing changes in such small increments that
commiting directly to the dev branch is feasible. The experience of individual
developers regarding version control is varied and some groups may benefit from
more advanced branching patterns that others.

\subsection{The change process}
The main idea behind configuration control is to achieve control of changes,
what impact they'll have and their costs.

Since this company is relatively small, the extent of CCB and it's usage will be
somewhat limited in contrary to the general idea of a CCB. The CCB consist of
the normal line up with: chairperson, secretary, members and specialists
\cite{daniels}. We will select the project manager to be the chairman and the
secretary will be the configuration manager in the company. The project manager
will appoint appropriate people to be members of the CCB and specialists will be
contacted on a ``when needed''-basis. There will likely be one representative
from each module with the exception that some modules representatives will be
considered experts instead, so that they will not be needed for each meeting.

The software in our case is as described earlier divided into modules. Each
module consists of around five developers and they're all located in the same
working area. The changes made on this level are traced through commit-comments
and the discussion is taken within the group orally.

The CCB are only involved when it comes to changes that cover more than one
module (the API between modules is modified). In the case that the CCB gets
involved, it will pretty much handle change requests that traditional way,
namely:

\begin{enumerate}
 \item A change is submitted to the SCM-manager.
 \item The SCM-manager decides if this change needs a meeting.
 \item If a meeting is needed, one is planned.
 \item On the meeting the chairman takes input and decide to approve or note
depending on the input.
\end{enumerate}

\subsection{The change process}
The main idea behind configuration control is to achieve control of changes,
what impact they'll have and their costs.

Since this company is relatively small, the extent of CCB and it's usage will be
somewhat limited in contrary to the general idea of a CCB. The CCB consist of
the normal line up with: chairperson, secretary, members and specialists
\cite{daniels}. We will select the project manager to be the chairman and the
secretary will be the configuration manager in the company. The project manager
will appoint appropriate people to be members of the CCB and specialists will be
contacted on a ``when needed''-basis. There will likely be one representative
from each module with the exception that some modules representatives will be
considered experts instead, so that they will not be needed for each meeting.

The software in our case is as described earlier divided into modules. Each
module consists of around five developers and they're all located in the same
working area. The changes made on this level are traced through commit-comments
and the discussion is taken within the group orally.

The CCB are only involved when it comes to changes that cover more than one
module (the API between modules is modified). In the case that the CCB gets
involved, it will pretty much handle change requests that traditional way,
namely:

\begin{enumerate}
 \item A change is submitted to the SCM-manager.
 \item The SCM-manager decides if this change needs a meeting.
 \item If a meeting is needed, one is planned.
 \item On the meeting the chairman takes input and decide to approve or note
depending on the input.
\end{enumerate}

The exception to the above rule are changes that are planned to already
released branches where the needs to be involved since the changes need more
rigorous control as they carry more risk not going through the normal
freeze/testing that predates a release.

\subsection{Bill of Material}

The company releases a new product every 6 months which results in several
different products to maintain. A problem is that when a customer finds a bug or
wants a new feature in a previous release The Problem Company can't reproduce
this system or it's very hard. We will need something so we both know exactly
what system the customer is using and so the programmer who will fix this
problem easy can reconstruct the system to work on. A way to do this is to use a
Bill of Material (BoM). When we release a product we then save information about
the release such as which version every component is using, what hardware are we
using, which features are implemented, test information etc. When a customer has
a problem with a certain release we can look at the BoM and know exactly what
system the customer is using and we can start working on a fix right away. We
will also need version control of the product to get the correct version of each
component in the program. Everyone in the project must be able to find this BoM
without too much effort and we need a system which makes it easy to setup a
certain BoM on the computer which the programmer is using. Otherwise we loose
the benefits with the BoM. The software which the company is producing are used
on specific cellphones so we need to be able to reproduce the hardware somehow.
This can be done by storing some cellphones but very often will the software not
interact with the actual hardware and we can then focus on just
software information. In the BoM we store both software information about our
product but also the software the cellphone is using.

\section{Tools}

\subsection{Version control}

\subsubsection{Considerations}
The branching patterns chosen for the company on the higher level don't have to
much of a requirement for advanced branch handling but the backporting of fixes
from the development branch to the release branches is a quite frequent
workflow and should be made easy. The releases need to be built from a
collection of releases and this could either be made with some kind script
checking out the specified versions of the modules or version of the modules
with the configurations and script verions controlled as well or as just the
build script with the modules assed to this project as submodules if the tool
has support for this. The latter is preferred if possible. As the developers
are mostly using eclipse as development platform integration with this is a big
advantage. The company will not outsource or create another division for the
forseeable future so a single central repository would work well and the
network can be utilized at LAN-speeds.

\subsubsection{Choices}
From the above requirements the choices are quite open. Most tools support the
branching required and the cherry-picking of changes can be made in most tools
albeit more or less easily. The distributed nature is not needed as access to
the server is good and most developers are not experienced enough to take
advantages of local commits and branches, and those that are can work that way
against the server using personal branches.

\bibliography{refs}
\bibliographystyle{plain}

\end{document}